---
title: "Analysis of simple Forex currency mean reversion"
author: "Emmanuel ROCHE"
date: "13/03/2016"
output: html_document
---

# Synopsis

In this analysis, we analyze the mean reversion property of pre-fabricated Forex currency pairs. We start with **USDCAD**, analyzing the daily close prices, from July 22, 2007 to March 28, 2012. This analysis is based on the original study provided in **Ernest P.Chan - Algoritmic trading winning strategies and their rationales**.


# Data preprocessing

We start with loading the appropriate dataset. We use the 1 minute data covering 01-2015 for this initial analysis:

```{r cache=TRUE}
#data <- read.csv("../RData/inputs/2015_01/USDCAD_M1_2015_01.csv")
data <- read.csv("USDCAD_D1_2007_07_2012_03.csv")

# assign appropriate names to this dataset columns:
names(data) <- c("date","time","open","high","low","close","volume")

# print some of the data:
head(data)
```


We have `r dim(data)[1]` observations available, but note that, this dataset we are only interested in the close prices for the time being, so we keep only that column:

```{r}
prices <- data$close

# let's now plot the data:

nob <- length(prices)
plot(1:nob,prices, type="l", col="blue", xlab="Time", ylab="Excahgne rate", main="USDCAD exchange rate")

```

# Results

## ADF test

Now we can perform the ADF test, for this we will use the tseries package, containing the **adf.test** function.

```{r}
library(tseries)

# With the default number of lag coefficients:
tres1 <- adf.test(prices)
tres1

# Now with only 1 lag coefficient:
tres2 <- adf.test(prices,"s",k=1)
tres2
```

In both test cases the p.values are quite high (respectively `r sprintf("%.3f",tres1$p.value)` and `r sprintf("%.3f",tres2$p.value)`) so we cannot reject the null hypothesis, and this currency is not stationary (which is of course expected).

## Hust exponent computation

We use the **pracma** package to compute the Hurst exponent. Note that we compute the Hurst exponent for the **log prices** serie.

```{r}
library(pracma)
hres1 <- hurstexp(log(prices))
```

The most "interesting value" in the previous list is the **Theoretical Hurst exponent** : `r sprintf("%.4f",hres1$Ht)`. Given this value is would seem that the currency pair analyzed in this period exibited a small trending tendancy.


## Variance ratio test

To check the statistical significance of this Hurst exponent value, we perform a **variance ratio** test. We will use the **vrtest** package to acheive this:

```{r}
library(vrtest)
nob <- length(prices)
lret <- log(prices[2:nob]) - log(prices[1:(nob-1)])
lres1 <- Lo.Mac(lret, kvec = 2)
lres1
```

The previous statistics indicate that we have `r sprintf("%.2f%%",-lres1$Stats[1]*100)` chances that the return values follow a random walk (??? Not really sure what those numbers mean in fact), so we cannot discard this hypothesis.

As described from [this page](http://stackoverflow.com/questions/14186547/vrtest-package-extended-text-answers) we can compute the p-value of this statistic usign the Boot.test function:

```{r}
Boot.test(lret, kvec=c(2,5), nboot=500,wild="Normal")
```

Right now, I cannot make any sense of those results unfortunately.

## Half-life of mean reversion


To determine the half-life $\lambda$ of our time serie we run a regression fit with $y(t) - y(t-1)$ as the dependent variable and $y(t-1)$ as the independent variable.

```{r}
nob <- length(prices)
xval <- prices[1:(nob-1)] 
dy <- prices[2:nob] - prices[1:(nob-1)]
reg <- lm(dy ~ xval)

plot(xval, dy, col='blue')
abline(reg,col='red')

# Now we can generate the value of the half-life:
lambda <- reg$coefficients[[2]]
halflife = -log(2)/lambda
```

So we find a $\lambda$ = `r sprintf("%.8f", lambda)` and the corresponding halflife is, HL = `r sprintf("%.0f days",halflife)`. This result is not exactly the same as what was mentioned in the original study (eg. 115 days) but in fact most of the previous results are somewhat different too.

## Simple linear Mean-Reverting strategy

Using the previous halflife value we compute the moving standard deviation of the price:

```{r}
# build an helper function to perform moving computations:
mov_apply <- function(x, n, callback){
    k = length(x);
    result = rep(NA, k);
    for(i in 1 : (k - n + 1)){
        result[i+n-1] <- callback(x[i :  (i + n - 1)]);
    }    
    return(result);
}

lag <- function(x,k)
{
  return(c(rep(NA,k),x[1:length(x)-k]))
}

movingAvg <- function(x,n)
{
  return (mov_apply(x,n,mean))
}

movingStd <- function(x,n)
{
  return (mov_apply(x,n,sd))
}

# Setting lookback to the halflife value:
lb <- halflife

# Compute the market value we should hold:
mktVal <- -(prices - movingAvg(prices, lb)) / movingStd(prices,lb);

# Compute the daily profit and loss value:
pnl <- lag(mktVal, 1) * (prices - lag(prices,1)) / lag(prices,1)

# head(pnl,200)
```

Now we can plot the cummulative profit and loss:

```{r}

# Remove the NA values and replace them with 0:
pnl[is.na(pnl)] <- 0

cumpnl <- cumsum(pnl)
nvals <- length(cumpnl)

plot(1:nvals, cumpnl, type='l', col='blue', xlab="Days", ylab="Cumulative P&L", main="Cummulative Profit and Loss")
```


The results observed here are highly similar to those presented in the original paper. Except that here we do not end with a positive profit value.



